---
output: html_document
---

## Comorbidity Visualizations

```{r}
library(dplyr)
library(ggplot2)
library(gtsummary)
```

```{r, here-i-am-comorbidities}

here::i_am(
  "comorbidities/report.Rmd"
)

data <- read.csv(
  file = here::here("data/covid_sub.csv")
)

```


```{r}
# Subset the dataset to include only relevant variables
comorbidity_data <- data %>% 
  select(X, AGE, SEX, PREGNANT, ICU, PATIENT_TYPE, COPD, DIABETES, DATE_DIED)

# Convert "Yes" to 1 and "No" or "N/A" to 0 for COPD and DIABETES variables
comorbidity_data <- comorbidity_data %>% 
  mutate(
    COPD = ifelse(COPD == "Yes", 1, 0),
    DIABETES = ifelse(DIABETES == "Yes", 1, 0)
  )
```

We visualize the prevalence of COPD within each age group. This helps demonstrate how COPD prevalence changes with age, providing insights into the relationship between age and COPD in COVID-19 patients.
```{r}

# Recategorize age variable
comorbidity_data <- comorbidity_data %>%
  mutate(age_group = case_when(
    AGE < 18 ~ "Children",
    AGE >= 18 & AGE <= 39 ~ "Young Adult",
    AGE >= 40 & AGE <= 59 ~ "Middle Adult",
    AGE >= 60 ~ "Old Age"
  ))


COPD_by_age <- comorbidity_data %>%
  filter(!is.na(age_group)) %>%
  group_by(age_group) %>%
  summarise(COPD_prevalence = mean(COPD == "1", na.rm = TRUE))


# Set the order of age groups
COPD_by_age$age_group <- factor(COPD_by_age$age_group, levels = c("Children", "Young Adult", "Middle Adult", "Old Age"))


# Visualize the relationship between age categories and COPD
ggplot(COPD_by_age, aes(x = age_group, y = COPD_prevalence)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Prevalence of COPD by Age Group",
       x = "Age Group",
       y = "COPD Prevalence") +
  theme_minimal()

# Save the COPD prevalence data as an RDS file
saveRDS(
  COPD_by_age,
  file = here::here("comorbidities/output/COPD_prevalence.rds")
)
```
COPD is more prevalent in older COVID-19 patients.


We use ICU admission as a measure of disease severity. We stratify by diabetes status to assess if there's a difference in ICU admission rates between diabetic and non-diabetic individuals with COVID-19. 
```{r}
# Create a summary table
diabetes_icu_table <- tbl_cross(
  data,
  row = DIABETES,
  col = ICU
)

# Add labels
diabetes_icu_table <- diabetes_icu_table %>%
  modify_header(update = list(label = "**Diabetes**")) %>%
  modify_spanning_header(update = list(label = "**ICU Admission**"))

# Print the formatted table
diabetes_icu_table


# Save the data as an RDS file
saveRDS(
  diabetes_icu_table,
  file = here::here("comorbidities/output/diabetes_ICU.rds")
)
```
