---
output: html_document
---

## Introduction

Coronavirus disease (caused by the COVID-19 virus) is an infectious disease and most
people infected with the COVID-19 virus will experience mild to moderate respiratory
illness and may recover without requiring special treatment. Vulnerable populations
including pregnant women, and older people are more likely to develop serious illness.
We will examine descriptive statistics for the population including age, sex, pregnancy, ICU admission, pneumonia, COPD, and diabetes. The selected variables are from a [larger COVID-19 database published by the Mexican government](https://datos.gob.mx/busca/dataset/informacion-referente-a-casos-covid-19-en-mexico).
We will create a logistic regression model for death using these variables.
Covariates can be included and excluded from the logistic regression model through the configuration file.


```{r, load-libraries, echo=FALSE}
library(labelled)
library(gtsummary)
library(dplyr)
library(here)
```


```{r, here-i-am-table1, echo=FALSE}
here::i_am(
  "table1/report.Rmd"
)

data <- read.csv(
  file = here::here("data/covid_sub.csv")
)

```


```{r, label-and-format, echo=FALSE}

#creating new variable called DEATH by recording DEAT_DIED variable
data$DEATH <- ifelse(is.na(data$DATE_DIED), 0, 1)


#labeling the study variables
var_label(data) <- list(
  AGE = "Age",
  SEX = "Sex",
  PREGNANT = "Pregnancy Status",
  ICU = "Patient admitted to ICU or not",
  PNEUMONIA = "Pneumonia",
  COPD = "Chronic Obstructive Pulmonary Disease",
  DIABETES = "Diabetes",
  DEATH = "Death of COVID-19 patient"
)

#Checking distribution for AGE variable: 71 NA's
# summary(data$AGE)

```

## Table 1

```{r, make-table, echo=FALSE}
##decriptive statistics table
data |>
  select("AGE", "SEX", "PREGNANT", "ICU", "PNEUMONIA", "COPD", "DIABETES", "DEATH") |>
  tbl_summary(by = "DEATH",
              type = list(AGE ~ 'continuous',
                          SEX ~ 'categorical',
                          PREGNANT ~ 'categorical',
                          ICU ~ 'categorical',
                          PNEUMONIA ~ 'categorical',
                          COPD ~ 'categorical',
                          DIABETES ~ 'categorical')) |>
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Descriptive Statistics by COVID-19 Death: 0=No, 1= Yes**") |>
  add_overall()|>
  add_p()

```


